---
title: "Cancer Statistic of Australia"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    source_code: embed
---

```{r setup, include=FALSE}
library(dplyr)
library(readr)
library(plotly)
library(flexdashboard)
```

```{r data-load}
cancer_incidence <- read_csv("Cancer-incidence-age-standardised-rates-5-year-age-groups.csv", 
                             skip = 5)

cancer_mortality <- read_csv("Cancer-mortality-age-standardised-rates-5-year-age-groups.csv", 
                             skip = 5)


filtered_incidence <- cancer_incidence %>% filter(`Age group (years)` == "All ages combined")
filterd_mortality <- cancer_mortality %>% filter(`Age group (years)` == "All ages combined")

#projected new case count
case_count_male <- filtered_incidence %>% 
                   filter(`Year` == "2021" & 
                          `Sex` == "Males" &
                          `Cancer group/site` == "All cancers combined")

case_count_female <- filtered_incidence %>% 
                     filter(`Year` == "2021" & 
                            `Sex` == "Females" &
                            `Cancer group/site` == "All cancers combined" )

#projected death
death_count <- filterd_mortality %>% 
               filter(`Year` == "2021" & 
                      `Sex` == "Persons" & 
                      `Cancer group/site` == "All cancers combined" )

#top 10 diagnosed Cancer based on ASR 
projected_incidence <- filtered_incidence %>% 
                       filter(`Year` == "2021" & 
                              `Sex` == "Persons" & 
                              `Cancer group/site` != "All blood cancers combined" & 
                              `Cancer group/site` != "All cancers combined" &
                              `Cancer group/site` != "All sarcomas combined" )


#top 10 Cancer based on mortality rate
projected_mortality <-  filterd_mortality %>% 
                        filter(`Year` == "2021" & 
                               `Sex` == "Persons" & 
                               `Cancer group/site` != "All blood cancers combined" & 
                               `Cancer group/site` != "All cancers combined" &
                               `Cancer group/site` != "All sarcomas combined" )


```

Dashboard
=======================================================================

Row 
-----------------------------------------------------------------------

### Total Number of Projected Death, 2021
```{r value-box1}
valueBox(death_count$Count,
         icon = "fa-skull",
         color = "#EB020A")
```

### Projected Number of New Cases, 2021
```{r value-box2}
valueBox(case_count_male$Count ,
         icon = "fa-male",
         color = "#3B9DFF")
```

### Projected Number of New Cases, 2021
```{r value-box3}
valueBox(case_count_female$Count,
         icon = "fa-female",
         color = "#EB44AE")
```

Row
-----------------------------------------------------------------------
### <b>Projected Cancer for 2021 with highest Incidence ASR</b>
```{r top-10-incidence}
projected_incidence$`Cancer group/site` <- factor(projected_incidence$`Cancer group/site`, 
                                                  levels = unique(projected_incidence$`Cancer group/site`)[tail(order(projected_incidence$`Age-specific rate (per 100,000)`),10)])

plot1 <- plot_ly(projected_incidence, 
                 y= ~`Cancer group/site`,
                 x=~`Age-specific rate (per 100,000)`) %>%
         add_bars(width = 0.5,
                 orientation = 'h') %>%
         layout(title = 'Projected Cancer for 2021, Incidence Rate',
                titlefont = list(size = 14),
                xaxis = list(title = 'Incidence ASR(per 100,000)'),
                yaxis = list(title = ""),
                legend = list(title=list(text='<b> Incidence Rate </b>')))

plot1

```

### <b>Projected Cancer for 2021 with highest Mortality ASR</b>
```{r top-10-mortality}
projected_mortality$`Cancer group/site` <- factor(projected_mortality$`Cancer group/site`, 
                                                  levels = unique(projected_mortality$`Cancer group/site`)[tail(order(projected_mortality$`Age-specific rate (per 100,000)`),10)])

plot2 <- plot_ly(projected_mortality, 
                 y = ~`Cancer group/site`,
                 x = ~`Age-specific rate (per 100,000)`) %>%
         add_bars(width = 0.5,
                 orientation = 'h') %>%
         layout(title = 'Projected Cancer for 2021, Mortality Rate',
                titlefont = list(size = 14),
                xaxis = list(title = 'Mortality ASR(per 100,000)'),
                yaxis = list(title = ""),
                legend = list(title=list(text='<b> Incidence Rate </b>')))
plot2 


```

Row
-----------------------------------------------------------------------
### Information
```{r info}
caption = "<b> Incidence : </b> New cases of Cancer reported. <br>
           <b> Mortality : </b> Death from Cancer. <br>
           <b> Age-standardised Rate (ASR) : </b> Rate of Cancer (cases diagnosed/ death) per 100,000. <br>"


valueBox("Information", 
         caption = caption, 
         icon ="fa-info-circle", 
         color = "#10b5b5")
```



Exploratory Graph 
=======================================================================

Sidebar {.sidebar data-width=240} 
-----------------------------------------------------------------------
<br>
**Select Cancer Type**
```{r}
selectInput("cancer_type", "Cancer Type", choices = unique(filtered_incidence$`Cancer group/site`), selected = "Breast cancer")

getCancerIncidence <- reactive({
  sprintf('Incedence ASR of %s by Sex', input$cancer_type)
})

getCancerMortality <- reactive({
  sprintf('Mortality ASR of %s by Sex', input$cancer_type)
})


```

**Select Year**
```{r}
selectInput("year", "Year", choices = unique(filtered_incidence$Year), selected = "2021")

getCancerIncidenceYear <- reactive({
  sprintf('Incedence ASR of %s by age-group for %s.', input$cancer_type, input$year)
})

getCancerMortalityYear <- reactive({
  sprintf('Mortality ASR of %s by age-group for %s.', input$cancer_type, input$year)
})

```


Row 
-----------------------------------------------------------------------
### <b> `r renderText(getCancerIncidence())` </b>
```{r}
renderPlotly({
  
plt <-  filtered_incidence %>% 
        filter(`Cancer group/site` == input$cancer_type) %>% 
        plot_ly( x = ~`Year`, 
                 y = ~`Age-specific rate (per 100,000)`,
                 color= ~`Sex`, 
                 colors = "YlOrRd", 
                 type = 'scatter', 
                 mode = 'lines+markers',
                 hovertemplate = paste('<b>Year </b>: %{x}<br>','<b>Rate </b>: %{y}')) %>%
        layout(title = 'Age-standardised Rates by Sex',
               xaxis = list(title = 'Year'),
               yaxis = list(title = 'ASR (per 100,000)'),
               legend = list(title=list(text='<b> Incidence Rate </b>')),
               hovermode = "x unified")

plt

})
```

### <b> `r renderText(getCancerMortality())` </b>
```{r}
renderPlotly({
  
plt2 <- filterd_mortality %>% 
        filter(`Cancer group/site` == input$cancer_type) %>%
        plot_ly( x = ~`Year`, 
                 y = ~`Age-specific rate (per 100,000)`,
                 color = ~`Sex`, 
                 colors = "Blues",
                 type = 'scatter', 
                 mode = 'lines+markers',
                 hovertemplate = paste('<b>Year </b>: %{x}<br>','<b>Rate </b>: %{y}')) %>%
        layout(title = 'Age-standardised Rates by Sex',
               xaxis = list(title = 'Year'),
               yaxis = list(title = 'ASR (per 100,000)'),
               legend = list(title=list(text='<b> Mortality Rate </b>')),
               hovermode = "x unified")

plt2

})
```

Row 
-----------------------------------------------------------------------

###  <b> `r renderText(getCancerIncidenceYear())`</b>
```{r}
renderPlotly({
  
plt3 <- cancer_incidence %>% 
        filter(`Cancer group/site` == input$cancer_type & `Year` == input$year) %>% 
        plot_ly(x = ~`Age group (years)`, 
                y = ~`Age-specific rate (per 100,000)`,
                color = ~Sex, 
                colors = "YlOrRd", 
                type = 'scatter', 
                mode = 'lines+markers',
                 hovertemplate = paste('<b>Age Group </b>: %{x}<br>','<b>Rate </b>: %{y}')) %>%
        layout(title = 'Age-standardised Rates by Sex',
               xaxis = list(title = 'Age Group'),
               yaxis = list(title = 'ASR (per 100,000)'),
               legend = list(title=list(text='<b> Incidence Rate </b>')),
               hovermode = "x unified")
  
plt3

})
```


###  <b> `r renderText(getCancerMortalityYear())`  </b>
```{r}
renderPlotly({
  
plt4 <- cancer_mortality %>% 
        filter(`Cancer group/site` == input$cancer_type & `Year` == input$year) %>%
        plot_ly(x = ~`Age group (years)`, 
                y = ~`Age-specific rate (per 100,000)`,
                color = ~Sex, 
                colors = "Blues", 
                type = 'scatter', 
                mode = 'lines+markers',
                 hovertemplate = paste('<b>Age Group </b>: %{x}<br>','<b>Rate </b>: %{y}'))%>%
        layout(title = 'Age-standardised Rates by Sex',
               xaxis = list(title = 'Age Group'),
               yaxis = list(title = 'ASR (per 100,000)'),
               legend = list(title=list(text='<b> Mortality Rate </b>')),
               hovermode = "x unified")
  
plt4

})

```



Reference
=======================================================================

Column
-----------------------------------------------------------------------

- Data Source: Australian Institute of Health and Welfare. 2021. Cancer Data - Australian Institute of Health and Welfare. [online] 
Available at: <https://www.aihw.gov.au/reports-data/health-conditions-disability-deaths/cancer/data> [Accessed 9 June 2021].

- Colorbrewer2.org. 2015. ColorBrewer: Color Advice for Maps. [online]
Available at: <https://colorbrewer2.org/> [Accessed 10 June 2021].

- Kulkarni, S., 2019. Building an HR Dashboard in R using Flexdashboard. [online] Medium. 
Available at: <https://towardsdatascience.com/building-an-hr-dashboard-in-r-using-flexdashboard-76d14ed3f32> [Accessed 11 June 2021].

- Plotly.com. 2019. Plotly R Graphing Library. [online]
Available at: <https://plotly.com/r/> [Accessed 11 June 2021].

- Voevodin, N., 2021. Maps | R, Not the Best Practices. [online] Bookdown.org. 
Available at: <https://bookdown.org/voevodin_nv/R_Not_the_Best_Practices/maps.html> [Accessed 11 June 2021].

- Berishvili, N., 2020. Create an Interactive Dashboard with Shiny, Flexdashboard, and Plotly. [online] Medium.
Available at: <https://towardsdatascience.com/create-an-interactive-dashboard-with-shiny-flexdashboard-and-plotly-b1f025aebc9c> [Accessed 11 June 2021].
